---
title: "exercise-11"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(skimr)

f <- "https://raw.githubusercontent.com/difiore/ada-datasets/main/Mammal_lifehistories_v2.txt"
d <- read_tsv(f, col_names = TRUE)


skim(d)

```

#There are 14 variables. Order, family, genus, and species are categorical variables.Mass, gestation, newborn, weaning, wean mass,AFR, ma,x.life, litter size, litter/year and refs are numeric variables/

**STEP 1**

```{r}

#install.packages("naniar")  
library(naniar)

d <- d |> replace_with_na_all(condition = ~ .x == -999)


colSums(is.na(d)) #check total missing values per variables


```

**STEP 2**

```{r}

d <- d |> select(-c(`litter size`, refs))

glimpse(d)

```

**STEP 3**

```{r}

dlog <- d |> 
  mutate(across(where(is.numeric), log))

head(dlog)

```

**STEP 4**

```{r}

dlog <- dlog |>mutate(
                 relGest = resid(lm(`gestation(mo)` ~ `mass(g)`, na.action = na.exclude)),
                 relWean = resid(lm(`weaning(mo)` ~ `mass(g)`, na.action = na.exclude)),
                 relAFR  = resid(lm(`AFR(mo)` ~ `mass(g)`, na.action = na.exclude)),
                 relLife = resid(lm(`max. life(mo)` ~ `mass(g)`, na.action = na.exclude)),
    
         relNewbornMass  = resid(lm(`newborn(g)` ~ `mass(g)`, na.action = na.exclude)),
         relWeaningMass  = resid(lm(`wean mass(g)` ~ `mass(g)`, na.action = na.exclude))
  )


head(dlog)
```

**STEP 5**

```{r}
library(ggplot2)
library(patchwork)

p1 <- ggplot(dlog, aes(x = order, y = relLife)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Residual Max Lifespan by Order", y = "relLife", x = "Order")

p1

p2 <- ggplot(dlog, aes(x = order, y = relNewbornMass)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Residual Newborn Mass by Order", y = "relNewbornMass", x = "Order")

p2

p3 <- ggplot(dlog, aes(x = order, y = relWeaningMass)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Residual Weaning Mass by Order", y = "relWeaning Mass", x = "Order")

p3

p1 + p2 + p3 + plot_layout(ncol = 1)
```

#Primates have the highest residual lifespan.

#Macroscelidea has the highest residual newborn mass.

#Perissodactyla has the highest residual weaning mass.

**Step 6**

```{r}
library(MuMIn)    
library(tidyverse)
#dropna from variables for model predictors
d_winnow <- c("gestation(mo)", "newborn(g)", "weaning(mo)",
                "wean mass(g)", "litters/year", "mass(g)")

#max.life nodel
d_life <- dlog |> drop_na(`max. life(mo)`, all_of(d_winnow))

head(d_life)

#AFR
d_afr <- dlog |> drop_na(`AFR(mo)`, all_of(d_winnow))

head(d_afr)


```

```{r}
#dataset for model

life_data <- d_life |> select(`max. life(mo)`, all_of(d_winnow))
afr_data  <- d_afr  |> select(`AFR(mo)`, all_of(d_winnow))

head(life_data)
head(afr_data)
```

```{r}

#fit full models

mod_life <- lm(`max. life(mo)` ~ ., data = life_data, na.action = "na.fail")
mod_afr  <- lm(`AFR(mo)` ~ ., data = afr_data, na.action = "na.fail")


```

```{r}
#dredge() for model selection
mods_life <- dredge(mod_life)
mods_afr  <- dredge(mod_afr)

head(mods_life)
head(mods_afr)
class(mods_life)
class(mods_afr)
  

#Best models

mods_life[1, ]
mods_afr[1, ]   



```

#For the response variable max.life, the best model based on AICc included four predictors: gst(mo) (β = 0.2941), ltt/yer (β = -0.2474), mss(g) (β = 0.0859), wnn(mo) (β = 0.1422).This model AICc is 264.6, (delta = 0, weight =1).

#For the response variable AFR(mo), the best model based on AICc included the following predictors: gst(mo) (β = 0.3515), ltt/yer (β = -0.7312), mss(g) (β = 0.0576), wnn(mo) (β = 0.1444).This model AICc is 412, (delta = 0, weight =1).

```{r}
#delta
sum(mods_life$delta < 4)  
sum(mods_afr$delta < 4)   
```

#For the max.life, there are 5 models with delta AICc of 4 or less.

#For the AFR(mo), there are 7 models with delta AICc of 4 or less.

```{r}
#extract top models

top_life <- get.models(mods_life, subset = delta < 4)
top_afr  <- get.models(mods_afr, subset = delta < 4)


```

```{r}
#model averagng

avg_life <- model.avg(top_life, fit = TRUE)
avg_afr  <- model.avg(top_afr, fit = TRUE)

summary(avg_life)
summary(avg_afr)

#CIA
confint(avg_life)
confint(avg_afr)

#variable importance
sw(avg_life)
sw(avg_afr)


```

#For max. life(mo), the variables gestation(mo), litters/year, and weaning(mo) appeared in all of the top models, each with a sum of weights equal to 1.00. The variable mass(g) was included in most top models (weight = 0.95), while newborn(g) and wean mass(g) appeared less frequently, with weights of 0.40 and 0.31, respectively.

#For AFR(mo), the variables gestation(mo), litters/year, and weaning(mo) also appeared in all of the top models, with a sum of weights of 1.00. The variable mass(g) appeared in over half of the top models (weight = 0.57), as did wean mass(g) (weight = 0.53), while newborn(g) appeared less frequently (weight = 0.35).

```{r}

plot(avg_life, full = TRUE, intercept = FALSE,
     main = "Model-Averaged Coefficients: Max Lifespan")

plot(avg_afr, full = TRUE, intercept = FALSE,
     main = "Model-Averaged Coefficients: AFR")


```
